### 2022-03-02 paper2appendix.tex

\newpage

\hypertarget{appendix-1-survey-questions}{%
\section{Appendix 1: Survey Questions}\label{appendix-1-survey-questions}}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Q181}: Now, we're going to transition a bit to ask you about the information you received during the Covid-19 pandemic.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Q170}: In the past 12 months, without searching for it, did you receive information about the Covid-19 vaccine from \ldots{} (check all sources you received information from)

\begin{itemize}
\tightlist
\item
  your doctor or other health professional? (1)
\item
  a person like friend, neighbor, or family member that you know ? (9)
\item
  From television news channel or the newspaper? (10)
\item
  From an online discussion group, listserve, or other online forum including a Facebook group or subreddit? (11)
\item
  From a social networking site such as Facebook timeline, Twitter status update, or LinkedIn? (12)
\item
  Other (13) \_\_\_\_\_\_\_
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Q171}: And how about you yourself intentionally looking for information about a Covid-19 vaccine? Such information can include things such as advice, clarification, facts, and experiences.

\begin{itemize}
\tightlist
\item
  Yes (4)
\item
  No (5)
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Q172}: How did you look for information about the Covid-19 vaccine?

\begin{itemize}
\tightlist
\item
  Asking a person like friend, neighbor, or family member that I know (1)
\item
  Asking my doctor or another health professional (4)
\item
  Posted queries in an online discussion group, listserve, or other online forum like a Facebook Group or Subreddit (5)
\item
  Posted queries on a social networking site such as Facebook timeline, Twitter status update, or LinkedIn (6)
\item
  Searched for my question using an online search engine such as Google or Bing (7)
\item
  Other (8) \_\_\_\_\_\_\_\_\_\_\_
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Q173}: What sort of information did you search for?
Separate different topics with a comma or semi-colon.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Q174}: How useful was the information you found?

\begin{itemize}
\tightlist
\item
  Extremely useful (22)
\item
  Very useful (23)
\item
  Moderately useful (24)
\item
  Slightly useful (25)
\item
  Not at all useful (26)
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Q175}: Did the information you learned affect your decision to get vaccinated against Covid-19?

\begin{itemize}
\tightlist
\item
  Yes (39)
\item
  No (40)
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Q179}: Did you receive a Covid-19 vaccine?

\begin{itemize}
\tightlist
\item
  Yes (9)
\item
  No (10)
\item
  I'm unsure or would not like to respond (11)
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Q180}: Do you plan to receive a vaccine for the prevention of the Covid-19 virus?

\begin{itemize}
\tightlist
\item
  Definitely not (9)
\item
  Probably not (10)
\item
  Might or might not (11)
\item
  Probably yes (12)
\item
  Definitely yes (13)
\item
  I would not like to respond (14)
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{gender}: What is your gender?

\begin{itemize}
\tightlist
\item
  Male (1)
\item
  Female (2)
\item
  Other (3) \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_
\item
  Prefer not to say (4)
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{hispanic}: Are you Hispanic, Latino/a/x, or Latin American Origin?

\begin{itemize}
\tightlist
\item
  Yes (1)
\item
  No (2)
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{race}: What is your race? If you are ``mixed race,'' select all that apply.

\begin{itemize}
\tightlist
\item
  White (1)
\item
  Black or African American (2)
\item
  American Indian or Alaskan Native (3)
\item
  Asian (please specify): (4) \_\_\_\_\_\_\_
\item
  Other (please specify): (5) \_\_\_\_\_\_\_
\item
  Prefer not to say (6)
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{educ}: What is the highest level of education that you have completed?

\begin{itemize}
\tightlist
\item
  Less than high school (1)
\item
  High school graduate (2)
\item
  Some college (3)
\item
  Associate's or Technical degree (4)
\item
  Bachelor's degree (5)
\item
  Graduate or professional degree (6)
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{income}: We would be interested to know roughly what your total household income
before taxes is. We mean income from all sources, including welfare, stock
dividends, other household members' income, etc. In 2020, in which bracket did your total family income fall?

\begin{itemize}
\tightlist
\item
  Under \$1,000 (1)
\item
  \$1,000 to \$9,999 (2)
\item
  \$10,000 to \$19,999 (3)
\item
  \$20,000 to \$29,999 (4)
\item
  \$30,000 to \$39,999 (5)
\item
  \$40,000 to \$49,999 (6)
\item
  \$50,000 to \$59,999 (7)
\item
  \$60,000 to \$74,999 (8)
\item
  \$75,000 to \$89,999 (9)
\item
  \$90,000 to \$109,999 (10)
\item
  \$110,000 to \$129,999 (11)
\item
  \$130,000 to \$149,999 (12)
\item
  \$150,000 or over (13)
\item
  Don't know (14)
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{age}: What is your age? (in years)