### 2022-03-02 tables/paper1/data-sources-table.tex

\begin{table}[!h]
\caption{\label{tab:data-sources-table}}
\centering
\begin{tabular}[t]{}
\end{tabular}
\end{table}