### 2022-03-02 tables/paper2/table-1-desc.tex

\begin{table}[!h]

\caption{\label{tab:table-1-desc}Descriptive Statistics for Dichotomous and Numeric Variables}
\centering
\begin{tabular}[t]{lrr}
\toprule
  & Mean & SD\\
\midrule
\cellcolor{gray!6}{Received info} & \cellcolor{gray!6}{\num{0.66}} & \cellcolor{gray!6}{\num{0.47}}\\
Received info, Doctor & \num{0.47} & \num{0.50}\\
\cellcolor{gray!6}{Received info, Person} & \cellcolor{gray!6}{\num{0.67}} & \cellcolor{gray!6}{\num{0.47}}\\
Received info, News & \num{0.65} & \num{0.48}\\
\cellcolor{gray!6}{Received info, Social Networking Site} & \cellcolor{gray!6}{\num{0.40}} & \cellcolor{gray!6}{\num{0.49}}\\
Received info, Online Forum & \num{0.35} & \num{0.48}\\
\cellcolor{gray!6}{Sought Info} & \cellcolor{gray!6}{\num{0.78}} & \cellcolor{gray!6}{\num{0.41}}\\
Sought Info, Doctor & \num{0.44} & \num{0.50}\\
\cellcolor{gray!6}{Sought Info, Person} & \cellcolor{gray!6}{\num{0.42}} & \cellcolor{gray!6}{\num{0.49}}\\
Sought Info, Social Networking Site & \num{0.18} & \num{0.39}\\
\cellcolor{gray!6}{Sought Info, Online Forum} & \cellcolor{gray!6}{\num{0.27}} & \cellcolor{gray!6}{\num{0.44}}\\
Sought Info, Online Search & \num{0.41} & \num{0.49}\\
\cellcolor{gray!6}{Vaccination Status} & \cellcolor{gray!6}{\num{0.84}} & \cellcolor{gray!6}{\num{0.36}}\\
Vaccination Outlook & \num{0.87} & \num{0.34}\\
\cellcolor{gray!6}{Age} & \cellcolor{gray!6}{\num{37.76}} & \cellcolor{gray!6}{\num{10.75}}\\
Hispanic or Latino/x & \num{0.15} & \num{0.36}\\
\cellcolor{gray!6}{Race, White} & \cellcolor{gray!6}{\num{0.88}} & \cellcolor{gray!6}{\num{0.33}}\\
Race, Black & \num{0.08} & \num{0.27}\\
\cellcolor{gray!6}{Race, Native American} & \cellcolor{gray!6}{\num{0.03}} & \cellcolor{gray!6}{\num{0.16}}\\
Race, Asian or Pacific Islander & \num{0.04} & \num{0.20}\\
\cellcolor{gray!6}{Associate's Deg or above} & \cellcolor{gray!6}{\num{0.74}} & \cellcolor{gray!6}{\num{0.44}}\\
\bottomrule
\multicolumn{3}{l}{\rule{0pt}{1em}Notes: 948 Surveyed, Conducted December 03 through December 12, 2021.}\\
\end{tabular}
\end{table}