### 2022-03-03 tables/paper1/data-sources-table.tex
fix paper 1 tables, finish most of writing

\newcolumntype{L}{>{\raggedright\arraybackslash}X}

\definecolor{Gray}{gray}{0.9}
\definecolor{lightblue}{HTML}{bad7db}

\renewcommand{\arraystretch}{1.2}

\begin{table}[!h]
\caption{\label{tab:data-sources-table}Data Sources}
\centering
\fontsize{8}{10}\selectfont
\setlength{\extrarowheight}{2pt}
\begin{tabularx}{\linewidth}{@{}L c c L@{}}
\toprule
\textbf{Validated Data Source} & \textbf{Type} & \textbf{Dates} & \textbf{Google Trends Used}\citep{googletrends} \\ \midrule
% \rowcolor{lightblue}
\multicolumn{4}{l}{\textbf{Behaviors and Attitudes}} \\\hline
% General Social Survey & Cross-Sectional & 2010 - 2020 & NA \\
Vaccine Hesitancy for COVID-19 \citep{vaches_data}& Cross-Sectional & March 3 - 15, 2021 & Search Topics: 'Covid-19 vaccine', 'Coronavirus (Disease)', 'Coronavirus   (Virus)', 'Vaccine' \\
Mask-Wearing Survey Data \citep{mask_data} & Cross-Sectional & July 2 - 14, 2020 & Search Topics: 'Coronavirus (Disease)', 'Coronavirus (Virus)', 'Cloth   Face Mask', 'Mask', 'Civil and Political Rights' \\
American National Election Survey \citep{anes_data}& Cross-Sectional & 2020 & NA \\
% \rowcolor{lightblue}
\multicolumn{4}{l}{\textbf{Health}} \\\hline
Covid Rates \citep{covid_data}& Longitudinal & Every Monday, 2020 - 2021 & 'Covid-19', 'Coronavirus', 'Taste Loss', 'Smell Loss' \\
Suicide Rates \citep{suic_data} & Longitudinal & Yearly 2010-2020 & Search Topics: 'Suicide', 'Depression', Search Term:' Suicide Hotline' \\
% \rowcolor{lightblue}
\multicolumn{4}{l}{\textbf{Political}} \\\hline
American National Election Survey \citep{anes_data}& Cross-Sectional & 2020 & NA \\
Presidential Election Results \citep{pres_data}& Cross Sectional & 2016 \& 2020 & Search Topics: 'Hilary Clinton', 'Donald Trump', 'Joe Biden' \\ \bottomrule
\end{tabularx}
\end{table}