### 2022-03-03 tables/paper2/model-2-phi.tex

\begin{table}[ht]

\caption{\label{tab:model-2-phi}Phi Coefficient ($\phi$) of binary association between each Information Search Vehicle}
\centering
\begin{tabular}{p{0.25\linewidth} p{0.10\linewidth} p{0.08\linewidth} p{0.12\linewidth} p{0.15\linewidth} p{0.12\linewidth}}
\hline
 & Person & Dr & Online Forum & Social Networking Site & Online Search\\
\hline
Person& NA& 0.28& 0.172& 0.275& 0.216\\

Dr & & & 0.192 & 0.152 & 0.098\\

Online Forum& NA& NA& NA& 0.371& -0.073\\

Social Networking Site & & & & & 0.057\\

Online Search& NA& NA& NA& NA& NA\\
\hline
\end{tabular}
\end{table}