### 2022-03-03 tables/paper2/table-1-desc.tex

\begin{table}[ht]

\caption{\label{tab:table-1-desc}Descriptive Statistics for Dichotomous and Numeric Variables}
\centering
\begin{tabular}{lrr}
\toprule
  & Mean & SD\\
\midrule
Received info & \num{0.66} & \num{0.47}\\
Received info, Doctor & \num{0.47} & \num{0.50}\\
Received info, Person & \num{0.67} & \num{0.47}\\
Received info, News & \num{0.65} & \num{0.48}\\
Received info, Social Networking Site & \num{0.40} & \num{0.49}\\
Received info, Online Forum & \num{0.35} & \num{0.48}\\
Sought Info & \num{0.78} & \num{0.41}\\
Sought Info, Doctor & \num{0.44} & \num{0.50}\\
Sought Info, Person & \num{0.42} & \num{0.49}\\
Sought Info, Social Networking Site & \num{0.18} & \num{0.39}\\
Sought Info, Online Forum & \num{0.27} & \num{0.44}\\
Sought Info, Online Search & \num{0.41} & \num{0.49}\\
Vaccination Status & \num{0.84} & \num{0.36}\\
Vaccination Outlook & \num{0.87} & \num{0.34}\\
Age & \num{37.76} & \num{10.75}\\
Hispanic or Latino/x & \num{0.15} & \num{0.36}\\
Race, White & \num{0.88} & \num{0.33}\\
Race, Black & \num{0.08} & \num{0.27}\\
Race, Native American & \num{0.03} & \num{0.16}\\
Race, Asian or Pacific Islander & \num{0.04} & \num{0.20}\\
Associate's Deg or above & \num{0.74} & \num{0.44}\\
\bottomrule
\multicolumn{3}{l}{\rule{0pt}{1em}Notes: 948 Surveyed, Conducted December 03 through December 12, 2021.}\\
\end{tabular}
\end{table}