### 2022-03-03 tables/paper2/table-2-desc.tex

\begin{table}[ht]

\caption{\label{tab:table-2-desc}Descriptive Statistics for Categorical Variables}
\centering
\begin{tabular}{llrr}
\toprule
  &    & N & Percent\\
\midrule
Gender & Female & 418 & \num{44.09}\\
 & Male & 524 & \num{55.27}\\
 & Other & 6 & \num{0.63}\\
Highest Education Level & Graduate or professional degree & 145 & \num{15.30}\\
 & Bachelor's degree & 560 & \num{59.07}\\
 & Associate's or Technical degree & 72 & \num{7.59}\\
 & Some college & 100 & \num{10.55}\\
 & High school graduate & 68 & \num{7.17}\\
 & Less than high school & 3 & \num{0.32}\\
Plan to be Vaccinated if not & Probably not & 38 & \num{4.01}\\
 & Definitely yes & 13 & \num{1.37}\\
 & Definitely not & 61 & \num{6.43}\\
 & Might or might not & 22 & \num{2.32}\\
 & Probably yes & 13 & \num{1.37}\\
Found Info Sought Useful & Very useful & 345 & \num{36.39}\\
 & Moderately useful & 112 & \num{11.81}\\
 & Extremely useful & 264 & \num{27.85}\\
 & Slightly useful & 18 & \num{1.90}\\
 & Not at all useful & 2 & \num{0.21}\\
Census Region & Midwest & 171 & \num{18.04}\\
 & Northeast & 170 & \num{17.93}\\
 & South & 365 & \num{38.50}\\
 & West & 223 & \num{23.52}\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}Notes: 948 Surveyed, Conducted December 03 through December 12, 2021.}\\
\end{tabular}
\end{table}