### 2022-03-05 tables/paper2/model-2-phi.tex

\begin{table}[ht]

\caption{\label{tab:model-2-phi}Phi Coefficient ($\phi$) of binary association between each Information Search Vehicle}
\centering
\begin{tabular}{p{0.25\linewidth} p{0.10\linewidth} p{0.08\linewidth} p{0.12\linewidth} p{0.15\linewidth} p{0.12\linewidth}}\toprule
 & Person & Dr & Online Forum & SNS & Online Search \\\toprule
Person & \- &  0.28& 0.172& 0.275& 0.216 \\
Dr & \- &  \- &  0.192 & 0.152 & 0.098 \\
Online Forum& \- &  \- &  \- &  0.371& -0.073 \\
SNS & \- &  \- &  \- &  \- &  0.057 \\
Online Search& \- &  \- &  \- &  \- &  \\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}Notes: SNS = Social Networking Site}\\

\end{tabular}
\end{table}