### 2022-03-15 tables/paper2/table-1-ind.tex

\begin{table}[!ht]

\caption{\label{tab:table-1-ind}Descriptive Statistics for Independent Variables}
\centering
\begin{tabular}{llrrr}
\toprule
  &    & Mean & SD & 95\% CI\\ \midrule
Sought Info & Overall  &  \num{0.78} & \num{0.41} & \num{0.76} to \num{0.81}\\
Sought Info & Doctor & \num{0.44} & \num{0.50} & \num{0.41} to \num{0.47}\\
Sought Info & Person & \num{0.42} & \num{0.49} & \num{0.39} to \num{0.45}\\
Sought Info & Online Search & \num{0.41} & \num{0.49} & \num{0.38} to \num{0.45}\\
Sought Info & Online Forum & \num{0.27} & \num{0.44} & \num{0.24} to \num{0.30}\\
Sought Info & Social Networking Site & \num{0.18} & \num{0.39} & \num{0.16} to \num{0.21}\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}Notes: 948 Surveyed, Conducted December 03 through December 12, 2021.}\\
\end{tabular}
\end{table}