### 2022-03-24 abstract.tex
review of ch 1

% at least 150 words, generally no more than 500 words

This dissertation focuses on how people search for information and how people
rely on this information to inform their health behaviors and develop social
norms. In academia and policy, the focus in research on 
information has most often focused on the information that is sent to 
consumers, seeing people as passive receivers of information. This
is otherwise known as a focus on the information 'push'. But, 
information is also intentionally sought out by individuals;
called information pull. My dissertation focuses on individuals as active agents 
in their search for information and how the information discovered through
these push and pull processes relates to the creation of social norms.

My first article tests the criterion validity of Google Search Trends as
an indicator of computer-mediated information search. I attempt to
validate Google Search Trends for use as indicators of attitudes, disease prevalence
and political preferences using five different validated data sources. 
My analysis revealed no correlation among any of the Google Trends tested and their
validated indicators. I demonstrate that there is no criterion validity of
Google Trends for the selected cases and social scientists will find no replacement for
high quality survey data with Google Trends. Instead, we must only use 
Google Trends to demonstrate interest or attention.

Knowing that Google Trends data only encompasses a small portion of the
information-seeking done by modern humans, my second article 
is motivated by the research question: How do computer-mediated or interpersonal 
information-seeking strategies vary across populations? Using original survey data
of 948 Americans, I investigate their experiences seeking out information about 
the COVID-19 vaccines. I investigate five distinct information seeking modalities, or methods of searching for information: 
personal connection, doctor, social networking site, online forum, and online search engine. 
% I find little evidence that online search is more utilized than
% seeking social support from personal network connections or health professionals
% as I had hypothesized based on uses and gratifications theory, 
% that media users are not passive consumers but instead actively choose
% media sources based on the satisfaction of their individual needs.
I find that different exposure points, the ways people first are exposed to 
information without searching for it, and information search modalities, the 
ways people intentionally search for information, hold real world consequences
through their associations with COVID-19 vaccination intentions and rates.
For example, receiving or seeking out information from a doctor increases 
 COVID-19 vaccination uptake while receiving information from a social 
 networking site was associated with lower odds of vaccination.

My final article takes a deeper dive into the formation of social norms
governing health behaviors in cases of extreme uncertainty, specifically using the cases of
 stay-at-home rates and vaccination rates as responses to public health
recommendations to mitigate the COVID-19 pandemic. Using the theories of associative diffusion
and the integrated theoretical framework of norms, I test models of behavioral adaption
to public health recommendations and patterns of complex contagion (the need for
repeated exposures to a something novel for it to diffuse) using linear
mixed effects models. My results show that complex contagion is a valid 
framework for the social contagion of new norms during COVID-19. 
However, I find an important novel moderating effect of signal discordance; 
if there is diversity in the information received by an ego, 
contagion is less likely to occur. This paper shows that
the contagion process is not fully understood without looking at the context
of each exposure to a contagion within the range of contagions one experiences.

The introductory chapter provides a summary of the research and an explanation of how this
research contributes to sociology, social science, and society. Namely, 
this paper provides important perspectives on search as an agentic process
and how the micro-level information seeking process of an individual can
lead to macro-level social norms. I show that information diffusion
is disrupted when conflicting information and behaviors are simultaneously
diffusing and therein contribute to research on diffusion, social networks,
and social norms.