### 2022-03-25 paper2.tex

\hypertarget{paper-2}{%
\chapter{information seeking Behaviors for Information on COVID-19 Vaccinations}\label{paper-2}}

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

Previous research has greatly failed to distinguish between the activation of
information seeking behaviors online and offline. Using theories of 1) social
support and 2) uses and gratifications theory I investigate the factors
associated with each type of information search: 
personal connection, doctor, social networking site, online forum, and online
search engine. Using novel survey data of 948 Americans and their experience
seeking out information about the COVID-19 vaccines, I find little evidence that
online search is more utilized than seeking social support from personal network
connections or health professionals. I find evidence that the modalities queried
in this survey are conceptually different and that the utilization of different
modalities varies by demography and information exposure points, where
individuals are provided with information without seeking it. Finally, I find that
different sources of information exposure and information search modalities hold
real world consequences through their associations with COVID-19 vaccination
intentions and rates, as information from a doctor increases COVID-19
vaccination uptake while receiving information from a social networking site
is associated with lower odds of vaccination.

\textbf{Keywords}: information seeking Behavior, COVID-19, Vaccine Hesitancy, Social Networks, Social Support, Information \& Communication Technologies

\hypertarget{intro-1}{%
\section{Introduction}\label{intro-1}}

Information is all around us and permeates every facet of human society, from
the hordes of data gathered during the internet age or passing gossip between
friends. Human communication and human society are based on the circulation of
information and knowledge. Information can be shared by others through
information campaigns and consumed by an individual, called information push, or
intentionally sought out by individuals, called information pull
\citep{cybenkoFoundationsInformationPush1999}.

Information is especially important given the prominent trends of
misinformation, disinformation, and mistrust in traditional institutions
\citep{starbird19, kata10}. While no belief in false information can be thought
of as harmless \citep{douglas21} some discredited theories such as 'vaccines
cause autism' have major ramifications for both the individual and public
health. For example, a newfound public resistance to vaccination against Measles
by so-called 'anti-vaxxers' caused a 17\% increase in rates worldwide in 2019.
This increase killed  142,000 people, most of whom were children under the age
of five \citep{givetash19}. However, even in the ever-evolving research agenda
of misinformation and infodemiology \citep{eysenbach02}, much of the research
focuses on the 'push' or 'supply' side of information, like spaces where
individuals first encounter with conspiracy theories or other information
\citep{johnsonOnlineCompetitionPro2020, broniatowski_etal20}.

The lack of research in which modalities are likely to be chosen when an
individual faces a need for information makes it difficult for policy makers to
propose effective information campaigns that promote the health and well-being
of individuals in our society. Even outside of the misinformation literature,
much of the research conducted in social networks and communication focus on
senders, influencers, and persuasion strategies
\citep{mertonManifestLatentFunctions1968, katzPersonalInfluencePart1955,
lazarsfeldPeopleChoice1944}, rather than on “the receiver as an active
information seeker and processor."
\citep{johnsonComprehensiveModelCancerRelated1993} \citep[for an exception,
see]{eysenbach09}.

I use the case of information around COVID-19 vaccinations to explore the
variation of usage of information search modalities. In doing so, I seek to
answer how computer-mediated or interpersonal information seeking strategies
vary across populations in their search for information about COVID-19
vaccinations and investigate how these strategies affect vaccination uptake.
Specifically, this article answers two research questions: (1) How do
computer-mediated or interpersonal information seeking strategies vary across
populations? (2) How does the information search modality utilized affect
COVID-19 Vaccination uptake?

Before addressing the results to these research questions, I offer an overview
of the current state of social science literature in regards to information
seeking strategies and providing a description of the research methodology and
analytic sample. I then discuss the implications of these findings for the
sociological understanding of information search and vaccination uptake.

\hypertarget{information-search}{\subsection{Information Search}\label{information-search}}

Active, directed searching by individuals to obtain information is sparsely
discussed in the sociological literature.
\citet{pejtersenDesignComputeraidedUsersystem1984}, a scholar of library and
information science, theorized that there are five strategies for searching for
information: browsing, analytical, empirical, known site, and similarity. The
most common strategy is browsing, where people follow leads based on
associations without premeditation. Another strategy is analytical, which
includes an explicit consideration of all facets of the question to guide a
search. The empirical method guides the search based on tactics that were
successful in past research. The known site strategy is to go to the direct
source of the information if known. And finally, the similarity method is to
find information based on another similar question that already has an answer.
These five strategies vary in their demands for prior knowledge, cognitive
processing, memory, and time spent. While this theory is aimed at finding
information in a library setting, scholars have extended the theory to other
fields and validated the framework \citep{fidelHumanInformationInteraction2012};
the frame is a useful beginning point for my theories of information search
through network activation or through computer-mediated communication.

Some communication theorists ask why an answer to a question is sought in the
first place. The Theory of Uncertainty Management
\citep{brashersCommunicationUncertaintyManagement2001} professes that people
search for information when their uncertainty around the subject leads to
anxiety or other cognitive harms. The Theory of Motivated Information Management
\citep{afifiTheoryMotivatedInformation2004, afifiSeekingInformationSexual2006}
extends the prior by adding that uncertainty itself is not the catalyst for
information-search; rather, it is driven by a discrepancy between the current
level of uncertainty on a subject and desired level of uncertainty. In other
words, individuals only perform information search when they are distressed by
their lack of knowledge or understanding on a subject.

One way to find information is to activate network ties to secure information
through a form of social support. Social support, while previously used
interchangeably with the terms social networks and social integration
\citep{houseStructuresProcessesSocial1988}, are the emotional, informational,
and instrumental assistance functions performed between social ties and have
strong and measurable association with health outcomes
\citep{houseMeasuresConceptsSocial1985, thoitsMechanismsLinkingSocial2011}.
Informational support is the process of seeking “help in defining,
understanding, and coping with problematic events and include education, advice,
or referral to another source of support” \citep[][p. 640]{winemiller_etal93}.
Brashers, a health communications researcher, defines informational support
slightly differently, focusing on the exchange of information that “facilitates
coping with life stresses… that may be exchanged among members of a support
network” \citet[][p. 260]{brashersInformationSeekingAvoiding2002}.These
definitions help to differentiate between seeking information through social
ties or through something anonymous like online search; a need for coping or
deeper understanding of important matters is thought to sway that decision.

Social support has long been theorized to stem from core discussion networks and
much of the historical social support surveys only examined these core networks.
The underlying assumption is that individuals reach out to a handful of strong
ties when in need of support, which could be elicited in surveys using name
generators \citep{marsdenCoreDiscussionNetworks1987}. This approach has yielded
important insights, but largely overlooks crucial processes of resource
activation \citep{hurlbertCoreNetworksTie2000, perrySocialNetworkActivation2015,
smithDonPutMy2005}. For instance, Small \citet{smallSomeoneTalk2017} shows that
the core discussion network does not capture how people activate social support
in practice and indicates that people draw on much broader social connections
for support, reminiscent of the weak ties research by Granovetter
\citet{granovetterStrengthWeakTies1973}. There is ongoing research investigating
how weak-tie support holds for different support types based on the architecture
put forward by \citet{houseStructuresProcessesSocial1988} of instrumental,
emotional, and informational support.

However, informational support can also be sought outside of the social network
context, namely via computer-mediated information search tools such as the
process of “Googling”, or the utilization of internet search engines. As the
online environment began penetrating all facets of modern human life, it makes
sense that performing online search has become one of the most convenient
modalities for information search. Small \citet{smallSomeoneTalk2017} showed
that Graduate students often confided in a "miscellaneous classmates encountered
down the hall" (p. 176) instead of established and trusted confidants because
they simply were there at the right time. I theorize that online search engines
may be the most frictionless and costless mode of support. No social capital or
relationship is necessary when choosing to utilize a search engine.

There are three computer-mediated modalities I consider for this analysis: online
search engines like Google or Bing, posting questions on online forums like
subreddits or Facebook groups, and posting a status update online via Facebook
or Twitter. While the first two are largely anonymous and do not require a
personal network, the third is a unique hybrid between personal network
connections and online search.

There are important factors that can influence information seeking strategies
besides the strength of social capital or ease of search. For instance, we can
borrow uses and gratifications theory (UGT) from studies of mass communication
\citep{blumlerUsesMassCommunications1974, tanMassCommunicationTheories1985}. UGT
posits that users are not passive consumers of media and that people have an
active role in choosing different sources of media based on their satisfaction
of specific needs on an individual basis. UGT is based on Maslow’s
\citeyear{maslowTheoryHumanMotivation1943} hierarchy of needs and is compatible
with Lazarsfeld and Katz’s theory of two-step flow
\cite{katzPersonalInfluencePart1955} because people can choose their media and
the opinion leaders they follow. Modern-day theorists have extended UGT theory
and classified the uses and gratifications of the internet and of social media.
\citet{staffordDeterminingUsesGratifications2004} theorize that the internet
provides gratification through useful content that meets expectations,
gratification from purposeful navigating or random browsing as a process, and
social gratification from forming and deepening social ties.
\citet{leungGenerationalDifferencesContent2013} theorizes that social media is
gratifying for users because it allows for venting of negative feelings,
provides recognition, provides entertainment, promotes social affection, and
fulfills cognitive needs.

\begin{hyp}[H\ref{hyp:online-vs-network}] \label{hyp:online-vs-network}
The proportion of survey respondents who seek information
using an online search engine will be significantly higher than those who
need to activate network resources, like asking a friend.
\end{hyp}

% this actually makes more sense not from UGT but as a transactional cost perspective (maybe like disposable ties), that you have to use social capital to activate your social network, but online search is free.

Adapting these facets of UGT and social support to my own case, I theorize that
online search allows for increased anonymity, a lowered social cost, and the
potential avoidance of embarrassment and other negative social interactions,
especially in the case on search engines.
\citet{rainsCopingIllnessDigitally2018} finds that patients tend to search for
technical information about an illness online but they then turn to their social
network for experiential information from others facing similar circumstances in
the case of cancer. The type of information sought is therefore important. If
specific expertise on a specific topic is desired by a searcher, they may reach
out to personal connections that they see as expert such as a doctor, or if an
individual distrusts the medical establishment, they may be more likely to
activate informational support among their social network or turn to online
groups that validate their worldview \citep{bogersHowSocialAre2014}. Finally,
UGT and Brasher's definition of social support
\citeyearpar{brashersInformationSeekingAvoiding2002} help us to theorize that
the psychosocial needs like identity management or relational maintenance of an
individual may influence the chosen information search modality: when the
informational need is related to forming and deepening social ties search is
likely to be done from network ties.

\begin{hyp}[H\ref{hyp:drgooglefriend}] \label{hyp:drgooglefriend}
Seeking information from experts, search engines, and other people
will be the most common modes of information search. 
\end{hyp}

\hypertarget{research-methodology}{\section{Research Methodology}\label{research-methodology}}

The data used for this research project are based on original survey data
collected between December 03, 2021 through December 12, 2021. This survey was
hosted on Qualtrics and participants were paid and recruited using Amazon Turk
(MTurk). Potential MTurkers were invited to participate if they were located in
the United States, had a task rate approval of over 95 percent, had completed 50
tasks, and had not completed this survey in previous sessions. The total valid
survey responses (\emph{n} = 948) were selected from a total of 1,066
respondents; some responses were disqualified due to a few factors such as
detected usage of a VPN, random answer clicking (determined by illogical
responses), poor quality typed responses (e.g., social network alters
consistently named random nouns), or taking the survey more than once. The
sample is slightly gender unbalanced, skewing towards male respondents. For the
respondents who provided their zip code (p = 0.98), the sample is quite balanced
with a slight skew towards MTurkers in the South (p = 0.41).

The survey took an average of 24.48 minutes with a standard deviation of 13.23
minutes. The shortest valid survey took 3.75 minutes and the longest took 111.08
minutes. Participants received \$6.00 in compensation for their participation in
the survey \footnote{All funding for this survey was provided by a Grant from
the Summer Institute in Computational Social Science and the Russell Sage
Foundation.}. The survey consisted of two main parts. The first portion of the
survey is a replication of they survey conducted by
\citet{smallSomeoneTalk2017}. In this section, we expand Small's research on
support activation by segregating weak-tie social support into instrumental,
emotional, and informational categories
\citep[see][]{houseStructuresProcessesSocial1988}. The second part of the survey
focuses on network resource activation and information seeking behaviors of
individuals regarding COVID-19 vaccinations for this paper.

\input{tables/paper2/table-1-ind}

\hypertarget{measures}{%
\subsection{Measures}\label{measures}}

\hypertarget{sought-information}{%
\subsubsection{Sought Information}\label{sought-information}}

There are multiple dependent variables of interest for this paper. The first
variable is a dichotomous indicator of whether someone intentionally sought out
information about COVID-19 vaccinations. The survey asked, ``And how about you
yourself intentionally looking for information about a COVID-19 vaccine? Such
information can include things such as advice, clarification, facts, and
experiences.'' About 78\% (\(\widehat{p}\))
of respondents had intentionally sought out information about COVID-19
vaccinations (95\% Confidence Interval = \{76\%, 81\%\}).

Further dependent variables segment the above question into multiple options;
``How did you look for information about the COVID-19 vaccine?'' Respondents
were posed with 5 responses and `other,' of which they could select multiple.
The largest proportion, 44\%, asked their `doctor or another health
professional'. About 42\% said they asked, `a person like friend, neighbor, or
family member that [they] know'. An additional 41\% searched `for [their]
question using an online search engine such as Google or Bing'. About a quarter
of the sample, 27\%, `posted queries in an online discussion group, listserv, or
other online forum like a Facebook Group or Subreddit'. And finally, 18\% of the
sample `posted queries on a social networking site such as Facebook timeline,
Twitter status update, or LinkedIn'. Much of the question wording and the
specific categories of search were inspired by ICPSR project 37220
\citep{scanlon19}.

\hypertarget{vaccination-outlook}{\subsubsection{Vaccination Outlook}\label{vaccination-outlook}}

\input{tables/paper2/table-2-dep}

Another Dependent variable for this analysis is a dichotomous variable
indicating whether the respondent has received or plans to receive a vaccination
against COVID-19. For this survey, the number of doses or timeline of receipt
were less relevant to the research question. This variable instead focuses on
intent or opinion about the vaccine. About 87\% have a positive view about the
vaccine. This construct is made through the combination of two survey items. The
first item is actual vaccination status, obtained through the question `Did you
receive a COVID-19 vaccine?' 84\% of the sample said that they had indeed
received a vaccination (95\% Confidence Interval = \{82\%, 87\%\}). As the
national vaccination rate is closer to 75\% \citep{cdc20}, this indicates some
bias in our sample that must be acknowledged: either MTurkers are more likely to
be vaccinated that the normal population (sampling bias) or there is some
conformity bias in the responses causing survey takers to provide false answers.
The second survey item that contributes to `Vaccination Outlook' was only given
to those who had not responded `yes' to the to their vaccination status. These
respondents were asked `Do you plan to receive a vaccine for the prevention of
the COVID-19 virus?' Of those who were asked this question (n = 147), 9\% said
they were `definitely yes' getting the vaccine; 9\% said they were `probably
yes' going to receive vaccine. The rest of the sample was more unsure: 15\% said
they `Might or might not,' 26\% said `Probably not,' and finally 41\% said they
would `definitely not' receive the vaccine. The variable `Vaccination Outlook'
indicates that either the respondent received a COVID-19 vaccine or either
probably or definitely will in the future.

\hypertarget{independent-variables}{%
\subsubsection{Independent Variables}\label{independent-variables}}

The first independent variables used in this analysis are based on the question,
`In the past 12 months, without searching for it, did you receive information
about the COVID-19 vaccine from \ldots{} (check all sources you received
information from).' This differs from the `How did you look for information'
question above because this is based on passive reception of information. The
largest proportion of respondents had received information from "a person like
friend, neighbor, or family member that [they] know" (\(\widehat{p}\) = 67\%).
Respondents also commonly received information from a television news channel or
a newspaper, about 65\% of the sample. 47\% of the sample had received
information from their doctor or other health professional, while 40\% received
information from a social networking site such as Facebook timeline, Twitter
status update, or LinkedIn. Finally, the lowest proportion of the sample had
received information from an online discussion group, listserv, or other online
forum like a Facebook group or subreddit (\(\widehat{p}\) = 35\%).

We also asked various demographic questions to describe our sample. The average
age of our sample was 37.76. 44\% of our sample self-identified as female, while
55\% self-identified as male. The sample is diverse racially though some ethnic
groups are underrepresented compared to national averages. Race was evaluated
using the question, `What is your race? If you are ``mixed race,'' select all
that apply.' Because of the multi-selection question, each variable is
dichotomous and proportions are shown. 88\% of the sample claimed they were
`White,' 8\% claimed to be `Black or African American,' 3\% chose `American
Indian or Alaskan Native,' 4\% selected `Asian,' and finally when asked, `Are
you Hispanic, Latino/a/x, or Latin American Origin?' 15\% selected `Yes.'
Respondents were also asked to select the highest level of education that you
have completed. Based on the breakdown between `Less than high school,' `High
school graduate,' `Some college,' `Associate's or Technical degree,' `Bachelor's
degree,' and `Graduate or professional degree,' 74\% of the sample were
classified to have a college-level education (Associate's degree or above).

\hypertarget{analysis}{%
\section{Analysis}\label{analysis}}

My analytic strategy proceeds in four steps. In Tables \ref{tab:table-1-ind} \&
\ref{tab:table-2-dep}, I present descriptive statistics for all study variables,
including variable ranges, means, and standard deviations. For this paper, I
rely heavily on logistic regression, modeled in \texttt{r}; specifically, I fit
a series of binomial generalized linear models with a logit link
\citep{venables2002a}, otherwise known as logistic regression model
\citep{hothorn2006handbook}. This approach is appropriate because the data are
non-nested, cross-sectional data with binary independent variables. In Table
\ref{tab:table-model-1} \footnote{All Tables in this paper were created with the
modelsummary package in \texttt{r} \citep{modelsummary}},I fit a logistic
regression model predicting whether someone sought information about COVID-19
using their sources of information (`Received info') as well as various
demographic variables such as age, education level and race. Figure
\ref{fig:plot-model-1} then illustrates the relationship between the
coefficients and information seeking.\footnote{Figures in this paper were all
created using the \texttt{r} packages jtools \citep{jtools} and ggplot2
\citep{wickham_etal, wickham11}}. In Table \ref{tab:model-2-phi}, I disaggregate
information seeking behaviors by investigating the difference in proportions of
each of the 5 different avenues of seeking information about COVID-19: from
another person, from a doctor, on an online forum, on a social networking site,
or using an online search tool. I investigate these differences through a phi
coefficient \citep{warrens08}, otherwise known as a mean square contingency
coefficient, which is used to investigate the degree of association between two
binary variables.\footnote{To calculate the Phi Coefficient, I utilize the Psych
\texttt{r} package \citep{psych}} I then investigate the associations between my
predictors and the five information search modalities by modelling a series of
logistic regressions in Table \ref{tab:table-model-2}. Finally, in Table
\ref{tab:table-model-3}, I investigate whether there is a relationship between
information exposure sources, information seeking behaviors and the choice to
get vaccinated against COVID-19 through a logistic regression model predicting
Vaccination Outlook. Figure \ref{fig:plot-model-3} then illustrates the
relationship between the coefficients and vaccination outlook.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{information seeking}{%
\subsection{Information Seeking}\label{information seeking}}

\input{tables/paper2/table-model-1}

\begin{figure}[h]
{\centering \includegraphics[width=0.8\linewidth]{figs/paper2/plot-model-1-1}}
\caption{What predicts someone intentionally searched for information? Plot of Coefficients}\label{fig:plot-model-1}
\end{figure}

Table \ref{tab:table-model-1} provides the results of my logistic regression
analysis predicting whether someone sought information about COVID-19. These
initial results suggest similar results for each modality of information
delivery: each mode of receiving information whether from a doctor, a person,
the news, a social networking site, or an online forum, all increase the odds of
seeking out information. For example, respondents who had received information
from a doctor or other medical professional were 2.82 (odds ratio = \(exp(\beta
k)\), \emph{p} \textless{} 0.001) times more likely to seek out information over
someone who had not. Similarly, respondents who had received information from `a
person like friend, neighbor, or family member that [they] know' were 1.87
(\emph{p} \textless{} 0.001) times more likely to seek out information over
someone who had not received information from a person like that. Interesting,
those who had received information from an online discussion group, listserv, or
other online forum including a Facebook group or subreddit were 2.39 (\emph{p}
\textless{} 0.001) times more likely to search out more information than someone
who had not. While most demographic variables do not seem to have a strong
effect on whether someone sought out information about the COVID-19 vaccine, a
few key variables had strong associations.  Respondents who are older are more
likely to seek information; The odds are about 3\% higher when age is equal to x
+ 1 relative to x, possibly indicating a generational effect of information
seeking patterns (\emph{p} \textless{} 0.01). Additionally, those who were more
educated, that is those with an Associate's degree or higher, were 1.58 more
likely to seek additional information than those without the degree (\emph{p}
\textless{} 0.05). And finally, there is evidence to suggest that the Black or
African American sample was more likely to seek out information, specifically:
Blacks were 3.58 times more likely to seek out information over non-Blacks
(\emph{p} \textless{} 0.05). Figure \ref{fig:plot-model-1} illustrates the
relationship between the coefficients and information seeking.

\hypertarget{information-search-modalities}{%
\subsection{Information Search Modalities}\label{information-search-modalities}}

\input{tables/paper2/model-2-phi}

To further investigate how people search for information, I asked those who had
sought information out intentionally, ``How did you look for information about
the COVID-19 vaccine?'' Because respondents could select as many responses as
applicable, each response is recorded as a dichotomous variable. Table
\ref{tab:table-1-ind} reveals differences in the prevalence of usage of the
various information seeking modalities: Around 40\% of the sample has inquired
for information from a doctor, a person, or using online search tool (44\%,
42\%, 41\%), while my respondents used modalities like online forums (27\%) and
social networking sites (18\%) to seek out information at significantly lower
rates.

To investigate how each Information Search modality are related to each other
using the Phi Coefficient ($\phi$) of binary association between each
Information Search modality pair. The Phi coefficient can be interpreted like a
correlation coefficient, with numbers closer to -1 indicating a very strong
negative relationship and those closer to positive 1 indicating a very strong
positive relationship (Warrens 2008). Table \ref{tab:model-2-phi} shows the
results of these coefficients. Out of the five information search modalities
investigated, there is evidence that the modalities are distinctly different
concepts because most coefficients fall under 0.3, the general cutoff for
correlation coefficients to indicate a relationship. There does seem to be a
relationship between Online Forums and social networking sites, possibly due to
an individual's latent propensity to use online networks over in-person
networks. These coefficients indicate a weak relationship between any two
modalities of information search and that respondents in my sample utilize each
of these a little bit differently.

I find that indeed the most common modalities are asking a doctor, asking a
friend, and asking a search engine which confirms Hypothesis
\ref{hyp:drgooglefriend}. However, because the modalities 95\% confidence
intervals in Table \ref{tab:table-1-ind} are overlapping, I fail to find
evidence for Hypothesis \ref{hyp:online-vs-network}, that using an online search
engine was more commonplace than asking a friend.

\input{tables/paper2/table-model-2}

I conduct a separate logistic regression for each search modality in Table
\ref{tab:table-model-2}. Modeling each response as a separate formula is useful
for interpretation, but there are important caveats: because each logistic
regression is on a different scale, one must not compare the magnitude of
coefficients between models. We can, however, use the direction and significance
of the coefficients to draw conclusions about these different forms of
information search modalities.

The first model in Table \ref{tab:table-model-2} predicts whether a respondent
sought information intentionally about COVID-19 vaccinations by asking their
doctor or another health professional. First and foremost, I find that those who
received information about COVID-19 vaccinations were 6.97 times more likely to
seek out more information from their doctor than those who had not (\emph{p}
\textless{} 0.001). However, receiving information from other modalities also
increased the odds of seeking out information from medical professions. Those
who received information from a person such as a friend or a relative were 2.23
times more likely to ask their doctor (\emph{p} \textless{} 0.001), while those
who received information from the news or television were 1.55 times more likely
(\emph{p} \textless{} 0.01). I also find that those who received information
from an online forum like a subreddit were 51\% more likely to ask their doctors
than those who had not (\emph{p} \textless{} 0.05). Finally, respondents with an
Associate's degree or higher were 72\% more likely to ask their doctor for
information on the COVID-19 vaccinations than those who held less education
(\emph{p} \textless{} 0.01).

The second model predicts whether a respondent sought information about COVID-19
vaccinations by asking a person like friend, neighbor, or family member that
they know. Only three coefficients in this model were significant and help us
draw any conclusions on who is most likely to activate their personal social
network for informational support. In this study, I find that receiving
information about the vaccine from a doctor, their personal network, or a social
networking site is associated with seeking out more information. Those who
received information from a doctor or other medical professional were 3.08 times
more likely to seek information from a person such as a friend (\emph{p}
\textless{} 0.001). Moreover, it seems that receiving information from your
personal social network, such as a friend or relative, is associated with
seeking information from those same people; in fact, I find that those who did
are 3.32 more likely to do so (\emph{p} \textless{} 0.001). Finally, the
respondents who received information about the vaccine on a social networking
site were 55\% more likely to seek information from their personal network than
those who had not (\emph{p} \textless{} 0.01). The unexpected result of this
model is that very few factors predict personal social network activation, and
this model has the lowest McFadden's R\^{}2, indicating possible omitted
variables.

I also wanted to examine what factors were associated with posting in an online
discussion group, listserv, or other online forum like a Facebook Group or
Subreddit about COVID-19 vaccinations. This is conceptually different than
posting a query on a social networking site like Facebook or twitter where
identities are involved, but is theoretically more social friction than
searching on an internet search engine. I found that having received information
about the vaccine from your personal social network (such as a friend or
relative), the news, or an online forum was associated with higher odds of
querying an online forum. Specifically, receiving information from your personal
network is associated with 2.07 higher odds (\emph{p} \textless{} 0.001), from
news or television media with 1.43 higher odds (\emph{p} \textless{} 0.05), and
receiving information on an online forum associated with 3.55 higher odds
(\emph{p} \textless{} 0.001). Moreover, I found that those with an Associate's
degree or above were 2.96 times more likely to search for information on an
online forum (\emph{p} \textless{} 0.001). Finally, I found that
Hispanic-Latinos were about 70\% more likely to query an online forum than
non-Hispanics (\emph{p} \textless{} 0.05). However, there were also a few
variables that lowered the odds of posting in an online discussion group. First,
those who had received information from a social networking site were 45\% less
likely to query a forum than those who had not (odds ratio = \(1 - exp(\beta
k)\), \emph{p} \textless{} 0.01). Moreover, I found additional ethnoracial
effects: Respondents who claimed Asian ancestry were 92\% less likely to query
online forum (\emph{p} \textless{} 0.01).

The fourth model predicts whether an individual sought information about
COVID-19 vaccinations through posted queries on a social networking site such as
Facebook timeline, Twitter status update, or LinkedIn. Receiving information
about COVID-19 seems to be quite related to posting on social media. Those who
received information from their doctors had 1.72 higher odds of posting on a
social networking site (\emph{p} \textless{} 0.01), while those who received
their information from tv/news had 1.67 higher odds (\emph{p} \textless{} 0.05).
Furthermore, those who received information on social networking sites were more
likely to post queries on social networking sites (\(\beta\) = 2.08, \emph{p}
\textless{} 0.001). The respondents in our sample who received information from
online forums had 2.65 higher odds of posting queries on their social networking
platforms (\emph{p} \textless{} 0.001). I found a relationship between education
and posting on social networking sites as well; specifically, those with an
Associate's degree or above were 3.83 times more likely to post vaccination
queries on social networking sites than those with lower education (\emph{p}
\textless{} 0.001). Finally, I find that Hispanic-Latinos were 88\% more likely
to ask their connections on a social networking site about the COVID-19
vaccinations than non-Hispanics (\emph{p} \textless{} 0.01).

Searching for information using an online search engine such as Google or Bing
is theoretically the search modality with the least amount of social or
cognitive friction. Our descriptive statistics showed that not everyone used
this modality, with only 41\% of our sample (See Table \ref{tab:table-1-ind})
having sought information about the COVID-19 vaccines using online search
engines. My model reveals interesting patterns to help explain the variation. As
with previous models, receiving information of any kind increase the odds of
online search, though to varying magnitudes (Doctor: 1.54 odds; Person: 1.44
odds; News or Television: 2.98 odds; social networking site: 3.57 odds; Online
Forum: 2 odds). I also find that older respondents were more likely to search
online, with each addition year aged to yield 3\% higher odds of searching
online (\emph{p} \textless{} 0.001). The most interesting variations come from
what decreases the odds of online search. First, those with an Associate's
degree or higher were 61\% less likely to search online for questions about
COVID-19 than those with less educational attainment (\emph{p} \textless{}
0.001). In addition, I find that Hispanic-Latinos had 44\% lower odds of
searching online than non-Hispanics (\emph{p} \textless{} 0.05).

\hypertarget{vaccination-views}{%
\subsection{Choosing to be Vaccinated}\label{vaccination-views}}

\input{tables/paper2/table-model-3}

Finally, in Table \ref{tab:table-model-3}, I investigate whether there is a
relationship between information receiving and information seeking behaviors and
the choice to get vaccinated against COVID-19 through a logistic regression
model predicting Vaccination Outlook. I find that receiving information from a
doctor or other medical profession is associated with 2.85 odds of choosing to
receive a COVID-19 vaccination (\emph{p} \textless{} 0.001) and that seeking
information from a doctor also increases the odds of vaccination by 4 (\emph{p}
\textless{} 0.001). I find that my respondents who received information on a
social networking site like Facebook or Twitter about the vaccines were 55\%
\textbf{less likely} to choose to get vaccinated than those who received no
information through that modality (\emph{p} \textless{} 0.001). However, I did
find that those who reached out to family or friends with their concerns were
77\% more likely to become vaccinated (\emph{p} \textless{} 0.05). Furthermore,
my respondents who held an Associate's degree or above had 3.6 higher odds of
becoming immunized (\emph{p} \textless{} 0.001). And finally, I found further
ethnoracial differences in my sample. I found that Hispanic-Latinos had 2.83
higher odds of vaccination than non-Hispanics (\emph{p} \textless{} 0.05) and
that my Asian sample had 4.73 higher odds as well over non-Asians (\emph{p}  =
0.09). See Figure \ref{fig:plot-model-3} for a visual representation of how
these different coefficients are related to vaccination outlook.

\begin{figure}[h]
{\centering \includegraphics[width=0.8\linewidth]{figs/paper2/plot-model-3-1}}
\caption{Does receiving or searching for information predict vaccination outlook? Plot of Coefficients}\label{fig:plot-model-3}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion and Conclusion}\label{discussion}}

Previous research in Sociology and Infodemiology has greatly failed to
distinguish between the activation of information seeking behaviors online and
offline. In this paper, I sought to uncover just how different various methods
of information search are and begin a line of research inquiry that investigates
the factors associated with each different modality. Finally, I also uncovered
how the different sources of information exposure and information search
modalities hold real world consequences through their associations with COVID-19
vaccination rates and intentions.

% TODO explain finding about online search being less prominent than social networks, though opposite to the theory. relate this to Rains work and how it may be the search for personalized or experiential information making the difference.

I first find that being exposed to information about COVID-19 vaccinations from
any source increases the likelihood of searching out more information. I also
find important demographic differences in the overall propensity to seek out
information. First, I find that the older a respondent is, the more likely they
are to seek out information about COVID-19 vaccinations, possibly indicating a
larger information need by elderly populations overall, or a more pressing need
to search for information about the COVID-19 vaccinations because of the higher
risk of hospitalizations that older populations face \citep{turner_etal18}.
Furthermore, I find that those who are college educated were more likely to seek
out information; for the most part this trend can be observed in Table
\ref{tab:table-model-2} as well. Human Capital Theory \citep{mirowsky_ross98}
may help to explain this finding; this theory claims that higher education can
provides the skills to gain health-related knowledge and use this knowledge to
be proactive in their own lives. Finally, I find that my respondents who
identified as Black were more likely to seek out further information; this may
be explained by the maltreatment that African Americans have endured in the US
healthcare system \citep{baileyStructuralRacismHealth2017} which has led to
major distrust of the medical establishment among the population
\citep{center2019, murray15, bronson2014don}.

Table \ref{tab:table-model-2} further demonstrated important takeaways about how
individuals seek information. First, I find that the different information
seeking modalities queried in this survey are conceptually different and that
the utilization of different varies by demographic and information exposure
source. Overall, being exposed to information about COVID-19 vaccinations from
any source is associated with information seeking behaviors, though there are
key variations. First, if one received information from a doctor, they were
likely to search from information from a doctor; this may indicate an already
existing trusting relationship between patient and healthcare provider but it
may also indicate the healthcare provider taking initiative and emphasizing the
importance of the vaccine or the survey respondent seeking out information from
an expert. Another important finding is that if you received information from a
social networking site like Facebook or Twitter, you were less likely to query
Online Forums like Reddit or Facebook Groups; this may be an example of Uses and
Gratification Theory \citep{blumlerUsesMassCommunications1974}, where the two
online platforms may act as substitute sources of information and those who were
satisfied with the information they received on a social networking site didn't
have the need to seek out more information. Finally, the ethnoracial variations
exhibited in Table \ref{tab:table-model-2} likely demonstrate how cultural
attitudes around the trustworthiness of different sources of information affect
the utilization of different methods; for example, Asians seem are less likely
to use online forums than non-Asians, while Hispanics are more likely to use
them than non-Hispanics. The internal variations between the different
information search modalities hint at different user profiles and may have major
repercussions on the quality of information they find and who is exposed to poor
quality information.

Given the rampant misinformation around COVID-19, its related vaccinations
\citep{pathakInfodemicsCOVID19Role2020, mottaHowRightLeaningMedia2020,
shahsavariConspiracyTimeCorona2020}, and the slowed rate of vaccinations in the
United States, it is important to uncover what relationship the different
sources of information exposure and information search modalities for COVID-19
vaccinations are associated with compliance with public health recommendations.
Table \ref{tab:table-model-3} demonstrated just how these different sources of
information exposure and information search modalities are associated with real
world consequences through COVID-19 vaccination rates and intentions. The first
takeaway is that receiving or seeking information from a doctor or other medical
profession was associated with higher likelihood of completed or planned
vaccinations. While blanket advice like ``Healthcare Providers should recommend
the vaccine to their patients'' would be ineffective because of the declining
number of Americans with a primary care doctor \citep{levine_etal20}, increasing
trust in the institution of medicine seems to remain a strong predictor of
vaccinations. A second major takeaway is that receiving information from a
Social Networking Site like Facebook timeline or Twitter status update was
significantly associated with lower vaccination rates. While we do not know what
exact information was absorbed on these platforms, the lower odds of vaccination
indicate that it was not positive. Much has been attempted between 2020 and 2022
to curb vaccine misinformation on these social networking platforms
\citep[see][for an early example]{bowman20}; however, I find evidence that these
social networking platforms nevertheless have a measurable negative impact.
Finally, this study also contributes to the conversation on ethnoracial and
educational differences in vaccine hesitancy. I find in my sample that
Hispanic-Latinos and Asians were more likely than non-Hispanics and non-Asians
to receive or plan to receive a vaccination against COVID-19 (Table
\ref{tab:table-model-3}). Research in this area has mixed results, with some
studies showcasing consistent findings to these
\citep{bagasra_etal21,king_etal21}, while others show that Whites have the
lowest levels of vaccine hesitancy and many minorities are much more hesitant
\citep{momplaisir_etal21, foxworth21}. My finding that college-educated
respondents were more likely seek vaccination is largely consistent with the
broader literature m\citep{khairat_etal22}, though again there are many
variations depending on the vaccine in question and study context
\citep{siddiqui_etal13}.

Previous research has greatly failed to distinguish between the activation of
information seeking behaviors online and offline and largely focuses on the
information that is sent to consumers [push], seeing people as passive receivers
of information. In this paper, I instead focus on individuals as active agents
in their search for information to begin a line of research inquiry that
investigates the factors associated with the various methods of information
search.

Given the large swaths of both misinformation and disinformation regarding
COVID-19 \citep{pathakInfodemicsCOVID19Role2020, mottaHowRightLeaningMedia2020,
shahsavariConspiracyTimeCorona2020} and the measurable impacts this
misinformation has had on pandemic-related health behaviors
\citep{loombaMeasuringImpactCOVID192021, greene_murphy21}, COVID-19 provides an
ideal study to investigate how we choose to search for information that affects
both our own lives and the lives of others.

This paper has provided several key findings that can be used to motivate policy
interventions for the COVID-19 vaccination campaigns and in the future for other
information campaigns. However, these findings are not to be taken out of the
context of their limitations. This paper is built off a small survey sample
(948) that was hosted online and anonymously through MTurk, a service that
provides people with micro-jobs. Because of the structure of the website,
respondents are incentivized to fill out surveys and other tasks as quick as
possible and there is no incentive for respondents to be honest in their
responses. For the same reason, there were many responses that were poor quality
and removed from the sample before analysis. However, each of these limitations
are common critiques of survey research as a whole. Even given these
limitations, the finding of this paper are interesting and timely contributions
to the literature.

Overall, this research project is important because the modalities used in
information search impacts the information and quality of information found. By
looking at how individuals search, I aimed to identified pain points and focus areas
for future interventions (receiving information online; ethnoracial variations) 
in the misinformation process.